The Emperor Has No Clothes

### 2011-03-30 Winds/papers/LBG_Lya/lya_lbg.tex
jxp
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5797 ac11df5e-a788-c23b-a8e1-c758c323924d

\newcommand{\tpk}{$\tau_{\rm pk}$}
\newcommand{\mtpk}{\tau_{\rm pk}}
\newcommand{\mdv}{\delta v}
\newcommand{\ewabs}{$W_{\rm a}$}
\newcommand{\ewe}{$W_{\rm e}$}
\newcommand{\rtt}{$r_{\tau = 0.2}$}
\newcommand{\mrtt}{r_{\tau = 0.2}}
\newcommand{\bturb}{$b_{\rm turb}$}
\newcommand{\mbturb}{b_{\rm turb}}
\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}$* \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}$* \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\feiis}{\ion{Fe}{2}$*$}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {\, {\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

@@

>>>>

\newcommand{\tpk}{$\tau_{\rm pk}$}
\newcommand{\mtpk}{\tau_{\rm pk}}
\newcommand{\mdv}{\delta v}
\newcommand{\ewabs}{$W_{\rm a}$}
\newcommand{\ewe}{$W_{\rm e}$}
\newcommand{\rtt}{$r_{\tau = 0.2}$}
\newcommand{\mrtt}{r_{\tau = 0.2}}
\newcommand{\bturb}{$b_{\rm turb}$}
\newcommand{\mbturb}{b_{\rm turb}}
\newcommand{\taud}{$\tau_{\rm dust}$}
\newcommand{\mtaud}{\tau_{\rm dust}}
\newcommand{\maconfig}{2p$^6$3s}
\newcommand{\mbconfig}{2p$^6$3p}
\newcommand{\aconfig}{a~$^6$D$^0$}
\newcommand{\zconfig}{z~$^6$D$^0$}
\newcommand{\mvr}{v_{\rm r}}
\newcommand{\naid}{\ion{Na}{1}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgii}{\ion{Mg}{2}}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\newcommand{\feiia}{\ion{Fe}{2}~$\lambda 2586$}
\newcommand{\feiib}{\ion{Fe}{2}~$\lambda 2600$}
\newcommand{\feiic}{\ion{Fe}{2}$* \; \lambda 2612$}
\newcommand{\feiie}{\ion{Fe}{2}$* \; \lambda 2626$}
\newcommand{\feiid}{\ion{Fe}{2}~$\lambda\lambda 2586, 2600$}
\newcommand{\feiis}{\ion{Fe}{2}$*$}
\newcommand{\nmg}{$n_{\rm Mg^+}$}
\newcommand{\mnmg}{n_{\rm Mg^+}}
\newcommand{\nfe}{$n_{\rm Fe^+}$}
\newcommand{\mnfe}{n_{\rm Fe^+}}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {\, {\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

@@

In a recent publication \citep[][hereafter PKR11]{pkr11}, 
we explored the absorption and emission from
a series of simple wind models with emphasis on the
\mgiid\ doublet.  In addition to a
series of idealized models, we also considered two variations of the
LBG wind model proposed by S+10.  Although we argued that the results
for \mgii\ should act as an analog for \lya\ because each is a
resonantly trapped doublet, we return in this paper to consider
directly ($\S$~\ref{sec:lya}).  We also expand upon our discussion on the required
energetics and mass flow of such a wind scenario
($\S$~\ref{sec:energy}).  Unless otherwise indicated, all distances
are physical under the assumption of a $\Lambda$CDM universe with
$\Lambda = 0.74$, $\Omega_m = 0.26$ and 
$H_0 = h_{72} 72 \, \mkms/{\rm Mpc}$ \citep{wmap05}.

\section{\lya\ Emission/Absorption from the LBG Clump Model}
\label{sec:lya}

The LBG clump model proposed by S+10 envisions an isotropic outflow of
dense clumps, each optically thick to \lya\ and metal-line
transitions, but as an ensemble have a partial covering fraction to
the LBG and background sources.  We discuss in PKR11 our radiative
transfer treatment for this wind in the context of \mgiid.  In the
following, we adopt the same prescription but consider
\lya\footnote{Following standard practice, we treat \lya\ as a single
  transition instead of a double because the line separation is much less
  than the astrophysical velocities of interest here.}
and modify the parameters of the LBG clump model to those presented in
S+10.

In contrast to PKR11, we include two additional features in the
model.  First, we have explored two intrinsic emission models for the
galaxy near the \lya\ line: 
(i) a flat continuum, primarily generated by O and B stars within the
star-forming galaxy; and
(ii) a continuum plus \lya\ emission-line model parameteized by a
line-width $\sigma_{\rm Ly\alpha}$ and an equivalent width \wlya.
This line is presumed to be generated by \ion{H}{2} regions within the
galaxy.  We emphasize that neithre of these simple models are likely
to describe the complex emission of a star-forming galaxy at \lya.
Nevertheless, they help capture some of the salient features of the
LBG clump model.

Second, we have modeled attenuation of photons near \lya\ by the
intergalactic medium (IGM) surrounding LBGs on scales of $\sim 100\,
{\rm kpc} - 10$\,Mpc.  Our approach follows the method of
\cite{santos04}, using the \cite{barkana04} model for infall onto
galactic halos at $z \approx 3$.  We assume a galaxy with dark matter
halo mass $M_{\rm DM} \approx 10^{12} \msun$, virial radius $\mrvir 
= 80$\,kpc, circular velocity $v_{\rm c} = 220 \mkms$, and an
IGM with $T = 10^4$\,K, and a EUVB with ionization rate $\Gamma =
10^{-12.5} \, \rm s^{-1}$ \citep{fpl+08}.  With this parameterization,
we can calculate the density and velocity profiles of gas at $r >
\mrvir$ for the LBG (and assume the gas at $r<\mrvir$ is dominated by
the outflow) and apply the Sobolev approximation to calculate the
optical depth \tauv\ to \lya\ as a function of velocity.  We cuation that
this $\exp(-\mtauv)$ prescription does not capture the complexity of
gas on these scales.  Indeed, \cite{zheng10a} have argued that a more
proper treatment tends to yield higher opacity.
[INCLUDE THE BOOST?]
In either case, it is a necessary consideration for modeling the
observed \lya\ emission from these galaxies.

Figure~\ref{fig:1D} (top panels) presents the 
calculated 1D \lya\ spectrum for the
two intrinsic emission models of the galaxy.
The second panels show the predicted results when one includes
radiative transfer through the LBG clump model.  Similar to the \mgii\
results of PKR11, we observe significant absorption by the outflow for
$v < 0\mkms$.  The absorption actually peaks at $v < 0\mkms$ because
scattered emission has `filled-in' the absorption for $v \approx 0
\mkms$.  This emission peaks just redward of $v=0\mkms$, related to
photons emitted transverse to our sightline and scattered once to
Earth, and then extends to positive velocity, with a tail to $v
\approx +500\mkms$.  [Comment on Lya emission line model?]
The third pair of panels includes the effects of IGM scattering.  For
the DM halo and IGM model assumed, we find a peak optical depth
$\mtmax \approx 7$ at $v \approx v_{\rm c}$ that rapidly declines to
$\mtauv < 0.5$ at $v=0\mkms$.  This implies complete attenuation of
the \lya\ emission at $v \approx v_{\rm c}$ but only modest
attenuation at lower velocity.  For this parameterization, it is evident that
the effects of the IGM cannot be ignored.  If the IGM were just $5
\times$ more neutral than modeled here, the \lya\ line would be
completely ``extinguished'' at $v \gtrsim 0 \mkms$, as predicted for
the $z \sim 6$ universe \citep{zheng+10a}.  By the same token, if the
galaxy emits ionizing photons then \tmax\ may fall below unity and the
IGM would play a minor role.  Clearly, a full treatment of these
effects is warranted for the $z \sim 3$ universe.

In the bottom panels, we present the full model convolved with an
ideal spectrometer with FWHM~$\approx 250\mkms$ compared against the
observed, average \lya\ emission profile of LBGs from
\cite{shapley03}.  
At this spectral resolution,
typical of LBG spectroscopic observations, the weak emission that was
apparent for the continuum model is lost and one predicts a simple
absorption profile that extends to positive velocity because of IGM
attenuation.  This model is a very poor description of the data.  
For the line+continuum intrinsic model, in contrast, it is difficult
to detect the outflow and the IGM attenuation is completely lost at
this spectral resolution within the \lya\ emission profile.
Neither model predicts a significant emission line that is shifted by
several hundred \kms\ redward of systemic.  Indeed, reviewing the
results on \mgii\ from PKR11, such a shift is predicted for any
outflow model that we have considered.  The only scenario in PKR11
that showed a significant shift was when we included significant
opacity at $v \approx 0 \mkms$ at the source, i.e.\ an ISM model.
This has the the effect of altering the intrinsic emission [How?] 
and when combined with an outflow can yield a redshifted emission
profile.  [Do this with Lya!] 
[Comment that infall can produce a redshift \lya\ line too, i.e. Zheng
et al.]

[Previous authors have modeled the \lya\ profiles of LBGs and have
concluded the redshift emission is well decribed by an outflow
\citep[e.g.][]{verhomme}.  Similar to the
ISM+wind model described above, these outflow models only predict a
redshift if there is a very large opacity at $v \approx 0\mkms$.  In
the \cite{verhomme} models, this is accomplished by considering a
thin expanding shell with a total \ion{H}{1} column density $\mnhi > 10^{20}
\cm{-2}$.  In this case, the damping wings of the gas imply a large
opacity at $v \approx 0 \mkms$ provided the shell has a speed less
than $\approx 500 \mkms$.  This model, then, predicts a very strong
absorption profile at $v < 0 \mkms$, which is generally not observed
in the LBG spectra.  One of the few exceptions is the galaxy cb58.]

Now consider the spatial distribution of \lya\ emission.
Figure~\ref{fig:SB} shows the predicted surface brightness profile
vs.\ projected radius for the flat-continuum intrinsic model and with
no IGM attenuation.  The signal is integrated from velocities $v =
[-100,400]\mkms$. Overplotted on the model, at an arbirtrary
normalization, is the fitted profile of S+11 to their observations,
$\mu(\rho) \propt \exp(-\rho/b)$ with $b=20$\,kpc.  In constrast to
the observations, the clump model predicts that XX\%\ of the
integrated flux is emitted at $\rho < 10$\,kpc and the model assumes
an exponential-like profile only at very large radii.  In short, a
proper radiative transfer treatment of the LBG clump model shows it
cannot reproduce the observed, extended \lya\ emission.

The explanation for the model's failure is twofold:  
(1) the covering fraction of the clumps decrease with radius to
$f_{\rm c} < 0.2$ at $\rho > 50$\,kpc; and
(2) the velocity gradient at $r>5$\,kpc is very shallow.  Therefore,
at large radii the outflow absorbs very few photons relative to the
inner regions.  In terms of the Sobolev approximation, we estimate
$\tau_{\rm S}(\rho = 60 \, {\rm kpc}) = XX \tau_{\rm S}(\rho=1\, {\rm
  kpc})$.  
Note that unlike the kinematic signatures, the inclusion of an ISM
component at small radii would not alter any of these conclusions.
[Add a last few lines]
